Article 1:	Construct Validity and Correspondence of Google Trends

### 2022-02-18 paper1-cleaning.Rmd
Done with wrangling more or less

# Functions
```{r}
find_mondays <- function(year) {
  mondays <- seq(ymd(glue("{year}-01-01")),ymd(glue("{year}-12-31")),by="1 day")
  mondays <- mondays[wday(mondays,label = TRUE) == "Mon" & mondays < today() - 7]
  return(mondays)
}
find_initial_trends <- function(keyword, period_list){
  period_list <- as.character(period_list)

@@

  return(trends)
}
find_scale_point <- function(keyword, year){
  if (year(today()) == year) {
    span <- glue("{year}-01-01 {as.character(today() - 7)}")
    } else {
      span <- glue("{year}-01-01 {year}-12-31")
    }

@@

  if (rescale %>% drop_na() %>% nrow() == 0) stop('no rows of data')
  return(rescale)
}
rescale_trends <- function(trends, rescale){
  rescale_ratio <- rescale %>%
    mutate(location = 'Los Angeles CA',
           hits_rescale = replace_na(hits_rescale, 0),
           hits_rescale = ifelse(hits_rescale == 0, 0.0001, hits_rescale)) %>% 
    left_join(trends %>% 
                 mutate(date = ymd(date)), by = c('date', 'location'))  %>% 
    mutate(rescale_ratio = hits_rescale/hits)  %>% 
    select(date, rescale_ratio)

@@

# attitudinal
• Vaccine Hesitancy
		○ https://data.cdc.gov/Vaccinations/Vaccine-Hesitancy-for-COVID-19-County-and-local-es/q9mh-h2tw/data

cross-sectional
dates - March 3, 2021 – March 15, 2021
```{r}
find_cross_sectional_trends('trend', '2021-03-03', 12)
```

• Mask Usage
		○ New York Times Survey
		○ https://github.com/nytimes/covid-19-data/tree/master/mask-use

cross-sectional
July 2 and July 14, 2020
```{r}
find_cross_sectional_trends('trend', '2020-07-02', 12)
```

@@

```{r}
covid_rates <- vroom('https://raw.githubusercontent.com/nytimes/covid-19-data/master/rolling-averages/us-counties-2020.csv') %>% bind_rows(vroom('https://raw.githubusercontent.com/nytimes/covid-19-data/master/rolling-averages/us-counties-2021.csv')) %>% 
  mutate(wday = lubridate::wday(date, label=TRUE),
         fips = str_sub(geoid, start = -5)) %>% 
  filter(wday == "Mon") %>% 
  select(fips, date, county, state, covid_rate = cases_avg_per_100k)

covid_gtrend <- find_year_trends('Covid-19', 2020) %>% 
  rowbind(find_year_trends('Covid-19', 2021))

find_year_trends('Cant smell', '2020-01-01', 365)
find_year_trends('Covid-19', '2021-01-01', 365)
```

@@

<!-- TODO need new wonder file, this one only has 1k counties??  -->

```{r suicide}
suicide_measures <- read_delim(here('data','suicide',
                                    'Multiple Cause of Death, 2018-2019, Single Race.txt')) %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  select(fips = county_code, county, deaths, population, crude_rate)

suicide_gtrends <- find_cross_sectional_trends(c('suicide', 'depression', 'suicide hotline'), 
                                               '2019-01-01', 365) %>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  mutate(location = str_to_lower(location))

suicide <- suicide_measures %>% 
  left_join(select(dma, FIPS, DMA_google), by = c('fips' = 'FIPS')) %>% 
  full_join(suicide_gtrends, by = c('DMA_google' = 'location'))

  select(dma = DMA_google,
         fips = county_fips, 
         clinton_p, `Hilary Clinton`, trump_p, `Donald Trump`)
```

# political
• Presidential results
https://github.com/tonmcg/US_County_Level_Election_Results_08-20
```{r pres_results}
pres_2016_results <- read_csv('https://raw.githubusercontent.com/tonmcg/US_County_Level_Election_Results_08-20/master/2016_US_County_Level_Presidential_Results.csv') %>% 
  select(county_name, county_fips = combined_fips, 
         totalvotes = total_votes, clinton = votes_dem, trump = votes_gop) %>% 
    mutate(county_fips = str_pad(county_fips, 5, pad = 0),
           clinton_p = clinton / totalvotes,
         trump_p = trump / totalvotes)

pres_2020_results <- read_csv('https://raw.githubusercontent.com/tonmcg/US_County_Level_Election_Results_08-20/master/2020_US_County_Level_Presidential_Results.csv') %>% 
  select(state = state_name, county_name, county_fips, 
         totalvotes = total_votes, biden = votes_dem, trump = votes_gop) %>% 
    mutate(biden_p = biden / totalvotes,
         trump_p = trump / totalvotes)
```
```{r pres_gtrend}
pres_2020_gtrend <- find_cross_sectional_trends(c('Joe Biden', 'Donald Trump'), '2020-01-01', 307) %>% 
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date)

pres_2016_gtrend <- find_cross_sectional_trends(c('Hilary Clinton', 'Donald Trump'), '2016-01-01', 307) %>% 
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date)
```
```{r}
pres_2016 <- pres_2016_results %>% 
  inner_join(select(dma, FIPS, DMA_google), by = c('county_fips' = 'FIPS')) %>%
  left_join(pres_2016_gtrend, by = c('DMA_google' = 'location')) %>%
  select(dma = DMA_google,
         fips = county_fips, 
         clinton_p, `Hilary Clinton`, trump_p, `Donald Trump`)

@@

>>>>

# Functions
```{r find_year_trends}
find_mondays <- function(year) {
  print('finding mondays')
  mondays <- seq(ymd(glue("{year}-01-01")),ymd(glue("{year}-12-31")),by="1 day")
  mondays <- mondays[wday(mondays,label = TRUE) == "Mon" & mondays < today() - 7]
  return(mondays)
}
find_initial_trends <- function(keyword, period_list){
  print(glue("find trends for '{keyword}'"))

  period_list <- as.character(period_list)

@@

  return(trends)
}
find_scale_point <- function(keyword, year){
  print('finding relevant scale point, using US-CA-803')
  if (year(today()) == year) {
    span <- glue("{year}-01-01 {as.character(today() - 7)}")
    } else {
      span <- glue("{year}-01-01 {year}-12-31")
    }

@@

  if (rescale %>% drop_na() %>% nrow() == 0) stop('no rows of data')
  return(rescale)
}
rescale_trends <- function(trends, rescale){
  print('rescaling trends')
  rescale_ratio <- rescale %>%
    mutate(location = 'Los Angeles CA',
           hits_rescale = replace_na(hits_rescale, 0),
           hits_rescale = ifelse(hits_rescale == 0, 0.0001, hits_rescale)) %>% 
    left_join(trends %>% 
                 mutate(date = ymd(date)), by = c('date', 'location'))  %>% 
    mutate(rescale_ratio = hits_rescale/hits)  %>% 
    select(date, rescale_ratio)

@@

# attitudinal

## Vaccine Hesitancy
		○ https://data.cdc.gov/Vaccinations/Vaccine-Hesitancy-for-COVID-19-County-and-local-es/q9mh-h2tw/data
		○ https://data.cdc.gov/Vaccinations/Vaccine-Hesitancy-for-COVID-19-County-and-local-es/q9mh-h2tw

cross-sectional
dates - March 3, 2021 – March 15, 2021
```{r vacc_hes}
vacc_hes_data <- vroom(here('data', 'vaccine-hesitancy', 'Vaccine_Hesitancy_for_COVID-19__County_and_local_estimates.csv')) %>% 
  janitor::clean_names() %>% 
  mutate(fips = str_pad(fips_code, 5, pad = '0')) %>% 
  select(fips, county_name, vacc_hesitant = estimated_hesitant_or_unsure)

vacc_hes_gtrends <- find_cross_sectional_trends(c('covid conspiracy',
                                                  'COVID-19 vaccine',
                                                  'Coronavirus',
                                                  'Covid-19'
                                                  # URLdecode('%2Fg%2F11j8_9sv06'),
                                                  # URLdecode('%2Fg%2F11j2cc_qll')
                                                  ), 
                            '2021-03-03', 12) %>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  janitor::clean_names()

vacc_hes <- vacc_hes_data %>% 
  left_join(select(dma, FIPS, DMA_google), by = c('fips' = 'FIPS')) %>% 
  full_join(vacc_hes_gtrends, by = c('DMA_google' = 'location'))

vacc_hes %>% 
  pivot_longer(cols = covid_conspiracy:covid_19) %>% 
  ggplot(aes(y = vacc_hesitant, x = value, color = name)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~ name)

write_rds(vacc_hes, here('data', 'cleaned_data', 'vacc_hes.rds'))
```

## Mask Usage
		○ New York Times Survey
		○ https://github.com/nytimes/covid-19-data/tree/master/mask-use

cross-sectional
July 2 and July 14, 2020
```{r Mask}

mask_data <- read_csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/mask-use/mask-use-by-county.csv') %>% 
  rowwise() %>% 
  mutate(mask_rare = NEVER + RARELY) %>% 
  select(fips = COUNTYFP, mask_rare)

mask_gtrends <- find_cross_sectional_trends(c('Face mask',
                                              'Mask',
                                              # 'Civil and political rights',
                                              'Cloth face mask'
                                              ),
                                            '2020-07-02', 12)%>%   
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date) %>% 
  janitor::clean_names()

mask <- mask_data %>% 
  left_join(select(dma, FIPS, DMA_google), by = c('fips' = 'FIPS')) %>% 
  full_join(mask_gtrends, by = c('DMA_google' = 'location'))

mask %>% 
  pivot_longer(cols = face_mask:civil_and_political_rights) %>% 
  ggplot(aes(y = mask_rare, x = value, color = name)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_wrap(~ name)

write_rds(mask, here('data', 'cleaned_data', 'mask.rds'))

@@

```{r covid}
covid_rates <- vroom('https://raw.githubusercontent.com/nytimes/covid-19-data/master/rolling-averages/us-counties-2020.csv') %>% bind_rows(vroom('https://raw.githubusercontent.com/nytimes/covid-19-data/master/rolling-averages/us-counties-2021.csv')) %>% 
  mutate(wday = lubridate::wday(date, label=TRUE),
         fips = str_sub(geoid, start = -5)) %>% 
  filter(wday == "Mon",
         county != 'Unknown',
         !(state %in% c('Puerto Rico', 'Virgin Islands', 'Northern Mariana Islands'))) %>% 
  select(fips, date, county, state, covid_rate = cases_avg_per_100k)

covid_gtrend_2020 <- find_year_trends('Covid-19', 2020) 
covid_gtrend_2021 <- find_year_trends('Covid-19', 2021) 
covid_gtrend  <- covid_gtrend_2020 %>% bind_rows(covid_gtrend_2021)

taste_gtrend_2020 <- find_year_trends('Taste Loss', 2020) 
taste_gtrend_2021 <- find_year_trends('Taste Loss', 2021) 
taste_gtrend  <- taste_gtrend_2020 %>% bind_rows(taste_gtrend_2021)

smell_gtrend_2020 <- find_year_trends('Smell Loss', 2020) 
smell_gtrend_2021 <- find_year_trends('Smell Loss', 2021) 
smell_gtrend  <- smell_gtrend_2020 %>% bind_rows(smell_gtrend_2021)

covid <- covid_rates %>% 
  inner_join(select(dma, FIPS, DMA_google), by = c('fips' = 'FIPS')) %>% 
  left_join(covid_gtrend, by = c('DMA_google', 'date')) %>% 
  left_join(taste_gtrend, by = c('DMA_google', 'date')) %>% 
  left_join(smell_gtrend, by = c('DMA_google', 'date')) %>% 
  janitor::clean_names()

covid %>% 
  filter(fips %in% c("08003", '55057', '51059', '39085', 
           # '36119',
           '34021', '12071', '04017')) %>% 
  pivot_longer(cols = c(covid_rate, covid_19, taste_loss, smell_loss)) %>% 
  ggplot(aes(x = date, y = value, group = name, color = name)) +
  geom_line() + 
  geom_point() +
  facet_wrap(~ glue('{county}, {state}'), scale = 'free_y') + 
  theme_minimal()

write_rds(covid, here('data', 'cleaned_data', 'covid.rds'))

@@

```{r suicide}
suicide_measures_both <- read_delim(here('data','suicide',
                                    'Multiple Cause of Death, 1999-2020.txt')) %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  select(fips = county_code, county, year = year_code, deaths_both = deaths, population)

suicide_measures_mcd <- read_delim(here('data','suicide',
                                    'Multiple Cause of Death, 1999-2020 only multiple cause.txt')) %>% 
  janitor::clean_names() %>% 
  filter(is.na(notes)) %>% 
  select(fips = county_code, county, year = year_code, deaths_mcd = deaths)

breaks <-  c(10,20,40,60,100,200,400,700,1300)
suicide_measures <- suicide_measures_both %>% 
  left_join(suicide_measures_mcd, by = c('fips', 'county', 'year')) %>% 
  mutate(deaths_mcd = na_if(deaths_mcd,'Missing'),
         deaths_both = na_if(deaths_both,'Missing'),
         deaths_mcd_supp = ifelse(deaths_mcd == 'Suppressed', '(0:9]', NA), 
         deaths_both_supp = ifelse(deaths_both == 'Suppressed', '(0:9]', NA), 
         deaths_mcd = na_if(deaths_mcd,'Suppressed'),
         deaths_both = na_if(deaths_both,'Suppressed'),
         deaths_mcd_cut = as.character(cut(as.numeric(deaths_mcd), breaks = breaks)),
         deaths_both_cut = as.character(cut(as.numeric(deaths_both), breaks = breaks)),
         deaths_mcd_cut = ifelse(!is.na(deaths_mcd_supp),
                                 deaths_mcd_supp,
                                 deaths_mcd_cut), 
         deaths_both_cut = ifelse(!is.na(deaths_both_supp),
                                  deaths_both_supp,
                                  deaths_both_cut)) %>% 
        select(fips, county, year, deaths_both_cut) %>% 
  mutate(deaths = factor(deaths_both_cut, ordered = T),
         deaths = fct_relevel(deaths, c("(0:9]",
                                        "(10,20]",
                                        "(20,40]", 
                                        "(40,60]",
                                        "(60,100]",    
                                        "(100,200]",
                                        "(200,400]",
                                        "(400,700]",   
                                        "(700,1.3e+03]")))
suicide_measures %>% 
  count(year, deaths) %>%
  ggplot(aes(x = deaths, y = n)) +
  geom_bar(stat = 'identity') +
  facet_wrap(~year)

append_years <- function(year_begin, year_end){
  years <- seq(year_begin,year_end,1)
  results <- tibble('location' = character(),
    'suicide' = numeric(), 
    'depression' = numeric(),
    'suicide hotline' = numeric(),
    'year' = numeric()
    )
  for (i in years){
    print(glue("finding trends for {i}..."))
    results <- find_cross_sectional_trends(c('suicide', 'depression', 'suicide hotline'),
                                glue("{i}-01-01"), 365) %>%
      pivot_wider(names_from = keyword, values_from = hits)  %>%
      select(-date) %>%
      mutate(year = i) %>% 
      bind_rows(results)
  }
  results
}

suicide_gtrends <- append_years(2010, 2020)

suicide <- suicide_measures %>% 
  left_join(select(dma, FIPS, DMA_google), by = c('fips' = 'FIPS')) %>% 
  full_join(suicide_gtrends, by = c('DMA_google' = 'location', 'year' = 'year'))  %>% 
  janitor::clean_names()

write_rds(suicide, here('data', 'cleaned_data', 'suicide.rds'))

suicide %>% 
  filter(fips %in% c("08003", '55057', '51059', '39085', 
           # '36119',
           '34021', '12071', '04017')) %>% 
  mutate(`actual deaths` = as.numeric(deaths) * 10) %>% 
  pivot_longer(cols = c('suicide', 'depression', 'suicide_hotline', 'actual deaths')) %>% 
  ggplot(aes(x = year, y = value, group = name, color = name)) +
  geom_line() + 
  geom_point() +
  facet_wrap(~ glue('{county}'), scale = 'free_y') + 
  theme_minimal()

# political
• Presidential results
https://github.com/tonmcg/US_County_Level_Election_Results_08-20
```{r pres}
pres_2016_results <- read_csv('https://raw.githubusercontent.com/tonmcg/US_County_Level_Election_Results_08-20/master/2016_US_County_Level_Presidential_Results.csv') %>% 
  select(county_name, county_fips = combined_fips, 
         totalvotes = total_votes, clinton = votes_dem, trump = votes_gop) %>% 
    mutate(county_fips = str_pad(county_fips, 5, pad = 0),
           clinton_p = clinton / totalvotes,
         trump_p = trump / totalvotes)

pres_2020_results <- read_csv('https://raw.githubusercontent.com/tonmcg/US_County_Level_Election_Results_08-20/master/2020_US_County_Level_Presidential_Results.csv') %>% 
  select(state = state_name, county_name, county_fips, 
         totalvotes = total_votes, biden = votes_dem, trump = votes_gop) %>% 
    mutate(biden_p = biden / totalvotes,
         trump_p = trump / totalvotes)

pres_2020_gtrend <- find_cross_sectional_trends(c('Joe Biden', 'Donald Trump'), '2020-01-01', 307) %>% 
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date)

pres_2016_gtrend <- find_cross_sectional_trends(c('Hilary Clinton', 'Donald Trump'), '2016-01-01', 307) %>% 
  pivot_wider(names_from = keyword, values_from = hits)  %>% 
  select(-date)

pres_2016 <- pres_2016_results %>% 
  inner_join(select(dma, FIPS, DMA_google), by = c('county_fips' = 'FIPS')) %>%
  left_join(pres_2016_gtrend, by = c('DMA_google' = 'location')) %>%
  select(dma = DMA_google,
         fips = county_fips, 
         clinton_p, `Hilary Clinton`, trump_p, `Donald Trump`)

@@

write_rds(pres_2016, here('data', 'cleaned_data', 'pres_2016.rds'))
write_rds(pres_2020, here('data', 'cleaned_data', 'pres_2020.rds'))

### 2022-02-18 paper1.Rmd
Done with wrangling more or less

---
title: "Article 1:	Information-Seeking Behaviors for Information on Covid-19 Vaccinations"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output:
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
  bookdown::pdf_document2:
    latex_engine: xelatex
    number_sections: true
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "My Library.bib"
indent: true
always_allow_html: true
---

@@

@@

# Intro 
New big data sources have led to vast possibilities for social science research
because they are bigger, cheaper, and already available  (King 2011; Lazer et
al. 2009; Salganik 2017). Before overenthusiastically embracing these sources
into our workflows, social scientists must clearly establish parameters under
which these data sources could be operationalized (Bail 2014; Lazer et al.
2014). As prior research outlined, the “quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data” (Lazer et al. 2014:1203). Building on
this prior research outlining these various issues with big data (boyd and
Crawford 2012; Lazer 2015), this paper tests the construct validity of Google
Search Trends as an indicator of three different cases, namely cultural
attitudes, disease prevalence and voting behavior. These three cases will be
tested using cultural indicators from the NORC General Social Survey, United
States county-level suicide rates from National Center for Health Statistics, US
cases rates of Covid-19 from The New York Times, historical US Presidential
Election results, and the American National Election Survey. Data will be
analyzed against corresponding operationalized Google Trends longitudinal data
using Pearson’s r for pairwise correlations, testing for the strength of
relationships between the Google Trend indicator and the respective comparison
indicator. This paper will contribute to the creation of methodological norms
and standards of how to use Google Trends as a big data source for societal
research and serve as a critical inquiry into the adoption of big data without a
critical eye for the ecological validity of the sources.

With the expansion of big data, some research has shown extremely innovative
methods that lead to groundbreaking results that are shown to be reliable. As an
example, Blumenstock et al. (2015) use county-level cell-phone records to
construct the distribution of poverty and wealth in Rwanda, a country where
national surveys and censes are rare and costly. However, Blumenstock et al.
(2015) go to great lengths to demonstrate that their operationalization of the
cell phone data creates a reliable and valid construct; few social science
papers utilizing big data dig into the construct validity of their metrics to
this extent and even fewer publications focus on methodological guidelines of
how to use sources of big data (Asseo et al. 2020; Stiles and Grogan-Myers
2018). However, research has shown the small adjustments to an algorithm or
metric may void any research insight we are able to pull from such data (Lazer
et al. 2014). Because of this, I propose a methodological validation study of
the Google Search Trends source of big data to investigate how it is advisable
to utilize this data in social scientific research.

I will use three categorizations of ways I propose Google Trends could be
operationalized for social scientific usage. First, I’ll test Google Trends as
an operationalization of cultural attitudes with the General Social Survey.
After Bail (2014)’s call for cultural sociologists to utilize the ever expanding
world of big data, Google trends as a data source began appearing in
sociological and social science research.  From research on mass shootings and
firearms (Brownstein, Nahari, and Reis 2020; Semenza and Bernau 2020), protest
and anti-Muslim sentiment (Bail, Merhout, and Ding 2018; Barrie 2020; Gross and
Mann 2017), to analyzing country-level changes in social perception (Reyes,
Majluf, and Ibáñez 2018), Google search trends are a new and innovative
indicator of cultural interest. Extending into social networks and culture,
Bail, Brown and Wimmer (2019) even used Google trends to measure how culture
spreads around the globe.

Google Search Trends have also been used continuously in estimations of disease
prevalence and population health in journals like the Journal of Medical
Internet Research. While much of this research has focused on the Covid-19
pandemic (Jimenez et al. 2020a, 2020b; Lim et al. 2020:20; Mavragani and Gkillas
2020; Nguyen et al. 2020; Todorova, Tsankova, and Ermenlieva 2021), other
research has investigated Google Trends as an indicator of wellbeing (Brodeur et
al. 2021; Carpi et al. 2020; Du et al. 2020), suicidality (Burnett, Eapen, and
Lin 2020), vaccination uptake (Dalum Hansen, Lioma, and Mølbak 2016), obesity
(Sarigul and Rui 2014), and even insomnia (Zitting et al. 2020), to cover a few
examples. For a partial review of other utilizations, see Nuti et al. (2014).
According to Jaidka et al (2021), the majority of studies profess a correlation
of > .70, “demonstrating the vast potential of Google Search as a proxy for
monitoring population health” (p. 3) based on assumptions that individuals
search because of self-diagnosis and to identify possible courses of treatment
(De Choudhury, Morris, and White 2014).

Various sources have also used Google Trends as a way to forecast political
elections and political attitudes (Wolf 2018). For instance, Swearingen and
Ripberger (2014) investigate how U.S. Senate Elections relate to attention
measured by search traffic. Prado-Román et al. (2020) compare how Google Search
trends are able to predict presidential election results in both the United
States and Canada. Finally, the OECD Development Centre is investigating how
Google data can help elucidate governments’ approval in Latin America (Montoya
et al. 2020)

@@

@@

The purpose of this paper is more methodological than theoretical, and I see
this paper having an impact on the social sciences and computational social
science as researchers pursue more projects using this source of big data.
Google Trends are relatively underutilized in the field compared to in the
health sciences and business. Once I assess how this data can be used, I would
like to be able to join Bail (2014) in encouraging social scientists to pursue
more research with big data while taking into account the potential pitfalls
with any source of big data (McFarland and McFarland 2015).

>>>>

---
title: "Article 1:	Construct Validity and Correspondence of Google Trends"
author: "Kelsey Gonzalez"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    number_sections: true
  bookdown::word_document2: 
    toc: true
    reference_docx: StylesTemplate.docx
header-includes:
   - \usepackage{siunitx}
   - \usepackage{setspace}
   - \usepackage{dcolumn}
   - \usepackage{caption}
   - \usepackage{longtable}
   - \usepackage{booktabs}
   - \doublespacing
   - \usepackage{placeins}
csl: american-sociological-association.csl
bibliography: "thesis.bib"
indent: true
always_allow_html: true
---

@@

Possible titles 
Digital Trace Data as Social Indicators: Indicators of Attention Rather Than Support

@@

# Intro 
New big data sources have led to vast possibilities for social science research
because they are bigger, cheaper, and already available
[@kingEnsuringDataRichFuture2011; @lazerComputationalSocialScience2009;
@salganikBitBitSocial2017]. Before overenthusiastically embracing these sources
into our workflows, social scientists must clearly establish parameters under
which these data sources could be operationalized
[@bailCulturalEnvironmentMeasuring2014; @lazerParableGoogleFlu2014]. As prior
research outlined, the “quantity of data does not mean that one can ignore
foundational issues of measurement and construct validity and reliability and
dependencies among data” [@lazerParableGoogleFlu2014, p. 1203]. Building on this
prior research outlining these various issues with big data
[@boydCriticalQuestionsBig2012; @lazerIssuesConstructValidity2015], this paper
tests the construct validity of Google Search Trends as an indicator of three
different cases, namely cultural attitudes, disease prevalence and voting
behavior. These three cases will be tested using cultural indicators from the
NORC General Social Survey, United States county-level suicide rates from
National Center for Health Statistics, US cases rates of Covid-19 from The New
York Times, historical US Presidential Election results, and the American
National Election Survey. Data will be analyzed against corresponding
operationalized Google Trends longitudinal data using Pearson’s r for pairwise
correlations, testing for the strength of relationships between the Google Trend
indicator and the respective comparison indicator. This paper will contribute to
the creation of methodological norms and standards of how to use Google Trends
as a big data source for societal research and serve as a critical inquiry into
the adoption of big data without a critical eye for the ecological validity of
the sources.

With the expansion of big data, some research has shown extremely innovative
methods that lead to groundbreaking results that are shown to be reliable. As an
example, @blumenstockPredictingPovertyWealth2015 use
county-level cell-phone records to construct the distribution of poverty and
wealth in Rwanda, a country where national surveys and censes are rare and
costly. However, @blumenstockPredictingPovertyWealth2015
go to great lengths to demonstrate that their operationalization of the cell
phone data creates a reliable and valid construct; few social science papers
utilizing big data dig into the construct validity of their metrics to this
extent and even fewer publications focus on methodological guidelines of how to
use sources of big data [@asseoTrackingCOVID19Using2020;
@stilesAssessingCriterionValidity2018]. However, research has shown the small
adjustments to an algorithm or metric may void any research insight we are able
to pull from such data [@lazerParableGoogleFlu2014]. Because of this, I propose
a methodological validation study of the Google Search Trends source of big data
to investigate how it is advisable to utilize this data in social scientific
research.

I will use three categorizations of ways I propose Google Trends could be
operationalized for social scientific usage. First, I’ll test Google Trends as
an operationalization of cultural attitudes with the General Social Survey.
After @bailCulturalEnvironmentMeasuring2014’s call for cultural
sociologists to utilize the ever-expanding world of big data, Google trends as a
data source began appearing in sociological and social science research.  From
research on mass shootings and firearms [@brownsteinInternetSearchPatterns2020;
@semenzaInformationseekingWakeTragedy2020], protest and anti-Muslim sentiment
[@bailUsingInternetSearch2018; @barrieSearchingRacismGeorge2020;
@grossThereFergusonEffect2017], to analyzing country-level changes in social
perception [@reyes_etal18], Google search trends are a new and innovative
indicator of cultural interest. Extending into social networks and culture,
@bailPrestigeProximityPrejudice2019 even used Google
trends to measure how culture spreads around the globe.

Google Search Trends have also been used continuously in estimations of disease
prevalence and population health in journals like the Journal of Medical
Internet Research. While much of this research has focused on the Covid-19
pandemic [@jimenez_etal20; @jimenezCOVID19SymptomGoogle2020;
@limEstimatingInformationSeekingBehaviour2020;
@mavraganiCOVID19PredictabilityUnited2020; @nguyenGoogleTrendsAnalysis2020;
@todorovaInternetBasedData2021] , other research has investigated Google Trends
as an indicator of wellbeing [@brodeurCOVID19LockdownsWellbeing2021;
@carpiTwitterSubjectiveWellBeing2020; @duCOVID19IncreasesOnline2020],
suicidality [@burnettTimeTrendsPublic2020], vaccination uptake
[@dalumhansenEnsembleLearnedVaccination2016], obesity
[@sarigulNowcastingObesityUsing2014], and even insomnia
[@zittingGoogleTrendsReveal2020], to cover a few examples. For a partial
review of other utilizations, see @nutiUseGoogleTrends2014. According to
@jaidkaInformationseekingVsSharing2021, the majority of studies profess a
correlation of > .70, “demonstrating the vast potential of Google Search as a
proxy for monitoring population health” (p. 3) based on assumptions that
individuals search because of self-diagnosis and to identify possible courses
of treatment [@dechoudhurySeekingSharingHealth2014].

<!-- Following reports say that Google searches can predict covid  -->
<!-- 14. Brunori, P. & Resce, G. Searching for the peak Google Trends and the Covid-19 outbreak in Italy. SERIES working papers N. 04/2020 -->
<!-- Lampos, V. et al. Tracking COVID-19 using online search. Preprint at https://arxiv.org/abs/2003.08086 (2020). -->
<!-- 16. Walker, A., Hopkins, C. & Surda, P. The use of google trends to investigate the loss of smell related searches during COVID‐19 -->
<!-- outbreak. Int. Forum Allergy Rh. 10, 839–847 (2020). -->
<!-- 17. Stephens-Davidowitz, S. Google Searches Can Help Us Find Emerging Covid-19 Outbreaks. The New York Times https://www. -->
<!-- nytimes.com/2020/04/05/opinion/coronavirus-google-searches.html (2020). -->
<!-- 18. Frank, T. Goldman says fewer ‘loss of smell’ Google queries suggest better COVID outlook. Consumer News and Business Channel. CNBC https://www.cnbc.com/2020/04/15/goldman-says-fewer-loss-of-smell-google-searches-is-a-positive-sign-for-pande -->
<!-- mic.html (2020) -->

Various sources have also used Google Trends as a way to forecast political
elections and political attitudes [@wolfTrendingRightDirection2018]. For
instance, @swearingenGoogleInsightsSenate2014 investigate how U.S. Senate
Elections relate to attention measured by search traffic.
@prado-romanGoogleTrendsPredictor2020 compare how Google Search trends are able
to predict presidential election results in both the United States and Canada.
Finally, the OECD Development Centre is investigating how Google data can help
elucidate governments’ approval in Latin America [@montoyaUsingGoogleData2020].

@@

https://journals.sagepub.com/doi/10.1177/0894439316631043
			§ What constructs might google trends capture and not capture well? 
Capture attention but not attitudes

Also Asseo - "we assumed that media coverage may potentially decouple the search
popularity from the number of cases, since searches would result not only from
self-symptoms, but also from interest elicited by media coverage."

@@

The purpose of this paper is more methodological than theoretical, and I see
this paper having an impact on the social sciences and computational social
science as researchers pursue more projects using this source of big data.
Google Trends are relatively underutilized in the field compared to in the
health sciences and business. Once I assess how this data can be used, I would
like to be able to join @bailCulturalEnvironmentMeasuring2014 in encouraging social scientists to pursue
more research with big data while taking into account the potential pitfalls
with any source of big data [@mcfarlandBigDataDanger2015].
